---
title: "TermProject_jMadsen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown

### STEP 1

## Problem
# Can a correlation between the Big Mac Index and the USA's past/current economic health be used
# to predict the future economic health in the USA?

## Why should someone care?
# With retail investors turning to apps in order to try and make profits in the stock market,
# looking at various data sets could help point a good time to purchase a larger investment/offload
# prior to a potential drop.

## Questions
# Which model should be used?
# Does the Big Mac Index accurately correlate to the USA economic health?
# What's the best way to clean the data and avoid excess noise?
# How do we take the model then apply future growth to obtain a guess?
# What is an acceptable value to indicate a correlation?
# Does the Big Mac Index drop during recessions?
# Does the Big Mac Index increase proportional to the growth of the GDP?

## Proposed course of attack
# First is to clean the data. I've grabbed global data, so creating subsets of USA data
# will be ideal. Then doing best fit lines for both the Big Mac Index and the Stock Index.
# Compare the lines to see if they're similar. There are vastly more data points in the stock index.
# Take whatever correlation to see if there are any insights the Indecies can provide and build models
# off the data. Append the growth value from the GEP data to these models to see if the correlation continues.

## How will this help?
# This process will determine if a correlation exists (part a of the problem), and then forecast out if the
# growth of the stock market/Big Mac Index are correlated (part b of the problem).

## location of data
# big_mac - https://github.com/TheEconomist/big-mac-data/blob/master/source-data/big-mac-source-data.csv
# Price of Big Macs over the last 20 years, as collected by the Economist around the world.

# gep_data - https://datacatalog.worldbank.org/search/dataset/0037888/Global-Economic-Prospects
# World Bank's semi-annual publication on the world economy, provides several years of forecasted growth

# stock_index - https://www.kaggle.com/mattiuzc/stock-exchange-data?select=indexProcessed.csv
# 14 stock exchanges since 1966 and their (almost) daily value.

## Packages
# ggplot2 definitely for starters.
# Other required packages will be loaded as needed. Need to remember to update this as I find more.

## Plots/Graphs
# plotting the Big Mac Index and Stock Market throughout the time period will be good to see
# a starting trend. Applying a best fit line and comparing the slope will also help.

## What don't I know
# How to take a model for one data set, then apply a differently formatted dataset to get modelled values?
# If the accuracy is low on the initial index models, will the forecasted growth matter?
# Does this growth account for inflation?

### STEP 2

## final dataset
# nya.subset is the bulkier dataset. I don't know how to best to tackle what dates matter the most.
# Everything else is contained enough to provide required information.

## What information is not self evident
# How does the cost of beef over the years impact the Big Mac index?
# Is this evident in the Stock Market Index?

## Different Ways
# I could take the monthly stock market data instead of daily.
# I could also focus the stock dates that closely match the Big Mac Index price days.
# Either of these would lower the fidelity of the model, since dips and recessions matter.

## Slice and dice
# Using subsets, I've dropped all the data that I deemed unnecessary.
# I've gutted the stock index data to only include the dates that the Big Mac file covers.

## Summarize
# Price fluctuations in the Big Mac Index should "almost" correlate to the same fluctations
# in the Stock Index. I can then prog that out to see what the future costs might be.

## Plots/Graphs
# plotting the Big Mac Index and Stock Market throughout the time period will be good to see
# a starting trend. Applying a best fit line and comparing the slope will also help.

## Machine Learning
# I plan on using logistic modeling. I don't know enough about ML to incorporate larger
# solutions to this data set.

## Questions
# Why male models?
# What can I do with this information?
# Which family of log modeling will be the way to go?


```{r}
# import the boys
library(ggplot2)

## Set the working directory to the root of your DSC 520 directory
setwd("C:\\Users\\MyDir\\Documents\\GitHub\\dsc520\\")

# pull in the data sets
big_mac <- read.csv("data/big-mac-source-data.csv")
gep_data <- read.csv("data/GEPData.csv")
stock_index <- read.csv("data/indexProcessed.csv")

# time for some spring cleaning. Focus on 2000 and beyond in the USA.
# clean out any unneeded columns
nya.subset <- subset(stock_index, Index == "NYA")
nya.subset <- subset(nya.subset, Date >= "2000-04-01")
nya.subset <- subset(nya.subset, Date <= "2021-01-01")
nya.subset <- subset(nya.subset, select = -c(Volume, Open, High, Low, Adj.Close, CloseUSD))

big_mac.subset <- subset(big_mac, iso_a3 == "USA")
big_mac.subset <- subset(big_mac.subset, date <= "2021-01-01")
big_mac.subset <- subset(big_mac.subset, select = -c(name, iso_a3, dollar_ex, GDP_dollar))

gep.subset <- subset(gep_data, Country.Code == "USA")
gep.subset <- subset(gep.subset, select = -c(ï..Country.Name, Indicator.Name, Indicator.Code, X))


# Let's provide headers to show the jist of the data
head(nya.subset)
head(big_mac.subset)
head(gep.subset)


```

